The \emph{method of lines} refers to the application of the \emph{finite difference} method to a
semi-discretization of an equation, i.e. the spatial domain has been discretized by the finite
element method and then the time domain is discretized using a finite difference method. This is the
preferable method for discretization of the time domain, due to its simplicity and the fact that the
time domain does not have geometry.

In particular, we may apply a a \emph{backward Euler} finite difference scheme to \eqref{eqn:QGEFE}
which would produce the following discretization of the streamfunction form of the QGE
\begin{equation}
  a_0(\psi_h^{n+1} - \psi_h^n, \chi_h) + k\, \left[a_1(\psi_h^n,\chi_h) + a_2(\psi_h^n,\psi_h^n,\chi_h)
      + a_3(\psi_h^n,\chi_h)\right] = k\, \ell^n(\chi_h),\quad \forall \, \chi_h \in X^h.
  \label{eqn:BEQGE}
\end{equation}
where $k$ is the time step, subscripts represent the discretization in spatial domains, and
superscripts represent the discretization in time domain. Additionally, fince the linear form
$\ell(\chi_h)$ is, in fact, time dependent we place a superscript to indicate the discretization of
the $\ell$ in the time domain.  The attractiveness of using this backward Euler scheme is that it is
explicit and therefore one need not solve the non-linear system that would arize from the
\emph{forward Euler} scheme. However, it is well known that the backward Euler scheme is not
a-stable and therefore one needs to be careful to satisfy a CFL type condition.

Here in lies the problem with using a high order finite element scheme, such as the Argyris element,
and the method of lines; the backward Euler scheme is first order in time whereas the Argyris
element is sixth order in the $L^2$ norm. Thus, we might expect a CFL condition of the form
\begin{equation*}
  \frac{k}{h^6} \le C
\end{equation*}
where $C$ is a dimensionless constant that does not depend on either $k$ or $h$. Thus, we might
expect to need quite a small time step, which may be unreasonably small for numerical simulations.
Therefore, either a higher order finite difference scheme will need to be used, or an explicit
finite difference scheme will need to be employed. 

While an implicit scheme might seem an attactive option, one must take into account the ammount of
time required to solve the nonlinear system at each time step. For larger spatial discretizations
the time to solve such nonlinear systems might be prohibitive. Therefore, higher order finite
difference schemes appear to be the more attractive approach.

However, there may be hope for the explicit finite element schemes after all. It might be possible
to apply the Two-level method (\autoref{alg:TwoLevel}) to the QGE at each time step. That is we
solve the nonlinear system on a coarse mesh and then use that solution to linearize the nonlinear
system on the fine mesh. This method may allow for much quicker calculations than might be expected
using just an explicit finite difference scheme alone. We expect that this method will produce good
results.
