With the continuous increase in computational power, complex mathematical models
are becoming more and more popular in the numerical simulation of oceanic and
atmospheric flows. For some geophysical flows in which computational efficiency
is of paramount importance, however, simplified mathematical models are central.
For example, the \emph{quasi-geostrophic equations (QGE)}, a standard
mathematical model for large scale oceanic and atmospheric flows
\cite{Cushman11,Majda,Pedlosky92,Vallis06}, are often used in climate modeling
\cite{Dijkstra05}.

The QGE are usually discretized in space by using the \emph{finite difference
method} (FDM) \cite{San11}. The \emph{finite element method} (FEM), however,
offers several advantages over the popular FDM, as outlined in \cite{Myers}: (i)
an easy treatment of complex boundaries, such as those of continents for the
ocean, or mountains for the atmosphere; (ii) an easy grid refinement to achieve
a high resolution in regions of interest \cite{Cascon}; (iii) a natural
treatment of boundary conditions; and (iv) a straightforward approach for the
treatment of multiply connected domains \cite{Myers}. Despite these advantages,
there are relatively few papers that consider the FEM applied to the QGE (see
for example: \cite{Cascon, Fix, LeProvost94, Myers, Stevens82}).

To our knowledge, \emph{all} the finitie element (FE) discretizations of the QGE have been
developed for the streamfunction-vorticity formulation, none using the
streamfunction formulation. The reason is simple: The streamfunction-vorticity
formulation yields a second order \emph{partial differential equation (PDE)},
whereas the streamfunction formulation yields a fourth order PDE. Thus, although
the streamfunction-vorticity formulation has two variables ($q$ and $\psi$) and
the streamfunction formulation has just one ($\psi$), the former is the
preferred formulation used in practical computations, since its conforming FE
discretization requires low-order ($C^0$) elements, whereas the latter requires
a high-order ($C^1$) FE discretization.

The streamfunction formulation is, from both mathematical and computational
points of view, completely different from the vorticity-streamfunction
formulation. Indeed, the FE discretization of the streamfunction formulation
generally requires the use of $C^1$ elements (for a conforming discretization),
which makes their implementation challenging. From a mathematical point of view,
however, the streamfunction formulation has the following significant advantage
over the vorticity-streamfunction formulation: there are optimal error estimates
for the FE discretization of the streamfunction formulation (see the error
estimate (13.5) and Table 13.1 in \cite{Gunzburger89}), whereas the available
error estimates for the vorticity-streamfunction formulation are suboptimal.

Despite the simplification made in formulating the QGE from the full-fledged
equations of ocean, the numerical simulation of the QGE is still computationally
challenging when integrating over long time periods, as is the case in climate
modeling. Therefore, it is necessary to reduce the computational cost of QGE
simulations. We will consider the \emph{Two-Level} method for reducing this
computational cost. To our knowledge it will be the first time this approach has
been applied to the streamfunction formulation of the QGE.

Two-level finite element discretizations are very promising approaches for
finite element discretizations of nonlinear partial differential equations
\cite{Fairag98,Layton93}. Two-level finite element discretizations aim to
solve a particular nonlinear elliptic equation by first solving the system on a
coarse mesh and then using the coarse mesh solution as a linearized variable to
solve the linearized elliptic equation on  a fine mesh. The attraction of such a
method is that one need only solve the nonlinear equations on a coarse mesh and
then use this solution to solve on a fine mesh, thereby reducing computational
time without sacrificing much in the way of solution accuracy. The development
of the two-level finite element discretization was originally used by Xu in
\cite{Xu94} and later algorithms for the Navier-Stokes Equations (NSE) were
developed by Layton \cite{Layton93}, Fairag \cite{Fairag98, Fairag03}, and Shao
\cite{Shao11}.

The goals of this paper are five-fold. First, we use a $C^1$ finite element (the
Argyris element) to discretize the streamfunction formulation of the QGE. To the
best of our knowledge, this is the \emph{first} time that a $C^1$ finite element
has been used in the numerical discretization of the QGE. Second, we derive
optimal error estimates for the FE discretization of the QGE and present
supporting numerical experiments. To the best of our knowledge, this is the
first time that optimal error estimates for the QGE have been derived.  Third,
we present a Two-Level algorithm for solving the streamfunction formulation of
the QGE and present the error analysis associated with this algorithm. To the
best of our knowledge, this is the \emph{first} time that such an algorithm has
been applied the the streamfunction formulation of the QGE.  Fourth, we will
present an LES closure model for the streamfunction formulation of the QGE and
its theoretical framework.
